---
title: "W01 - Leaflet Map" 
author: "Mia Thuge Kuntz"
date: "`r Sys.Date()`"
output: 
  rmdformats::readthedown:
    highlight: kate
---
# Assignment 1 - Leaflet Map
I have chosen the first option, where I am tasked with creating a standalone .html map of the points of different ruins discovered in the Blue Mountains in New South Wales.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r install, include=FALSE}
library(leaflet)
library(tidyverse)
library(here)
```

## The data
I first read in the data set provided by Adela for this course. It can be found via following link, but will otherwise be available in my Github repository for this assignment: https://github.com/CDS-AU-DK/cds-spatial/tree/main/Week01

```{r read data}
rcf18 <- read.csv("RCFeature2018.csv")
glimpse(rcf18)
```

The glimpse function gives me an overview of the different columns and data types in the data set. 

## Creating the Leaflet
To create a Leaflet showing the different points of data in a map I make use of the `leaflet` pipe function. In this pipeline I add the `addTiles` function along with the `addMarkers` function. It is in the markers function that I include the longitude and latitude columns of the data set, which makes the plotting of the ruins and their coordinates possible. I also add a popup feature, which adds the desired attributes to the markers on the map - that is the FeatureID, FeatureType and Description columns. 
```{r specify data}
leaflet() %>% 
  addTiles() %>% 
  addMarkers(lng = rcf18$Longitude, 
             lat = rcf18$Latitude,
             popup = paste0(rcf18$FeatureID, "<br>", 
                            rcf18$FeatureType, "<br>",
                            rcf18$Description))
```


I have now successfully created the map. As can be viewed, almost all of the markers are in the same area, and the viewer may feel the need to zoom in quite a lot before experiencing the markers individually.

A warning message appears when running the code, which warns me of a row with missing lat/lon values. A quick glimpse at my data set shows my, that the missing lat/lon values comes from the last row in the data set, which does not contain any information besides its FeatureID and FeatureTimestamp. I move forward by removing this row from my data set.
```{r}
rcf18 <- rcf18[-257,]
```


## Minimap and Measure
The assignments suggest adding a minimap and a way of measuring the dimensions of the map to my code. Both are great suggestions, as they each add a new feature to the map, which makes it easier for the user to keep themselves orientated to their position or to help interpret what they see.

To add the minimap feature, I make use of the `addMiniMap` function in my pipeline. To make it possible to measure distances and areas, I use the `addMeasure` function, where my specifications to the functions are based on previous experience of what works well, as well as what is considered the "standard" European measuring units. 
```{r minimap and measure}
leaflet() %>% 
  addTiles() %>% 
  addMiniMap(position = "bottomright") %>%
  addMarkers(lng = rcf18$Longitude, 
             lat = rcf18$Latitude,
             popup = paste0(rcf18$FeatureID, "<br>", 
                            rcf18$FeatureType, "<br>",
                            rcf18$Description)) %>% 
    addMeasure(
    position = "bottomleft",
    primaryLengthUnit = "meters",
    primaryAreaUnit = "sqmeters",
    activeColor = "#3D535D",
    completedColor = "#7D4479")
```

The map now contains a minimap as well as a way of measuring distances and areas.

## Differentiating the markers
Next part of the assignment is trying to differentiate the markers. This is, in this example, an excellent idea, since most of the markers are located near or almost on top of each other when viewed in the map. Unfortunately, I wasn't able to achieve this, and I therefor look forward to the solution, as I can definitly see the merits of applying this to the map.

## Clustering the markers
Last part of the assignment was adding clusters to the map. 
```{r cluster}
leaflet() %>% 
  addTiles() %>% 
  addMiniMap(position = "bottomright") %>%
  addMarkers(lng = rcf18$Longitude, 
             lat = rcf18$Latitude,
             popup = paste0(rcf18$FeatureID, "<br>", 
                            rcf18$FeatureType, "<br>",
                            rcf18$Description),
             clusterOptions = markerClusterOptions()) %>% 
    addMeasure(
    position = "bottomleft",
    primaryLengthUnit = "meters",
    primaryAreaUnit = "sqmeters",
    activeColor = "#3D535D",
    completedColor = "#7D4479")
```

From a far away view is would not be recommended to add the clustering to this particular map, as most of the data points are concentrated in the same place/area. When zooming in it can be argued for, but in the overall picture it would be recommended to leave the marker clusters out. 

## Creating a standalone .html map
Since adding clustering to the markers was not a recommended tool, when saving my map as a standalone .html file I choose to save the one without the clustering, and instead save the one with the added minimap and measuring tool. In order to do so, I first need to save my map as an object, since this is needed when running the `saveWidget` function to create the .html file.
```{r creating object}
rcf18map <- leaflet() %>% 
  addTiles() %>% 
  addMiniMap(position = "bottomright") %>%
  addMarkers(lng = rcf18$Longitude, 
             lat = rcf18$Latitude,
             popup = paste0(rcf18$FeatureID, "<br>", 
                            rcf18$FeatureType, "<br>",
                            rcf18$Description)) %>% 
    addMeasure(
    position = "bottomleft",
    primaryLengthUnit = "meters",
    primaryAreaUnit = "sqmeters",
    activeColor = "#3D535D",
    completedColor = "#7D4479")
```

```{r html}
library(htmlwidgets)
saveWidget(rcf18map, "RCFeature2018.html", selfcontained = TRUE)
```

The .html file is now saved as a standalone map, and can be viewed independently from the above code. Please note that the result can be found in my Github repository along with the .html file and .Rmd file for the whole RMarkdown code.





